### 2017-01-29 paper/rebuttal.txt

TARGET AUDIENCE

Several reviewers asked who the intended users of this tool are and
when the use case was compelling.  Briefly, the technology is aimed at
programmers.  If one were to ask random programmers whether they were
familiar with and knew how to use regular expressions, most would say
yes.  If one were to random programmers whether they were familiar
with lenses 100% would say no.  The central idea is to exploit
something that most programmers know well (regular expressions) to
help them write significant programs in a language that they do not
know well (lenses).  Programmers are often reluctant to learn new
things.  Our system helps them avoid that.

Several reviewers wondered whether regular expressions were really
necessary at all and whether we could get away without requiring them.
One of the reasons we feel our work is valuable is that it illustrates
an important practical tradeoff: by providing data formats, we can
infer transformations over significantly more sophisticated string
data types than related systems which do not use such information.
For instance, FlashFill cannot infer nested iterations.  17 of our
benchmarks have nested iterations and the deepest nesting structure
involved 4 levels of nesting.  As another example, FlashExtract does
not have unions, but it can represent them indirectly by using
options.  Still, in our tests (e.g. on the BibTeX data source),
FlashExtract could not infer transformations between string data types
that contain unions.  By the way, both FlashExtract and FlashFill are
wonderful systems, but they are optimized for other problems.  Our
work demonstrates the tradeoff between designs.  We hope to make this
point more clearly in a revised version of the paper and to quantify
the number of benchmarks that we handle that cannot be handled by
other string synthesis systems.

Several reviewers wondered whether lenses really were tricky to write
and suggested we show a concrete example (thank you -- we should
have).  The following lens translates between idealized BibTeX and
EndNote author formats (it was auto-generated by our tool):

@@

  let bibtex_to_endnote_au
    : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
    del "author={"
      . ins "au - "
      . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
      . (del " and "
           . ins "\n au - "
           . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
      . del "}"

Writing this lens from scratch is quite tricky and fiddly.  The
trickiest part is the nested swaps, which are fiddly to think about
and get right.  We would also argue that the cognitive model overall
is tricky and quite different from what many programmers are used to
--- one has to think in two directions at once.  Importantly, this is
an incredibly simple example -- Augeas and Boomerang users regularly
work with lenses significantly larger than this.  For example,
Boomerang's full version of an idealized BibTeX converter is over 300
lines of code, and all but 4 of the Augeas programs we studied were
over 50 lines of code, with 4 over 100 lines of code.

@@

FORMALIZATION

A second major concern was the lack of a formal treatment of our
implementation, in particular the function RigidSynth.

In earlier drafts, we actually had a formal description of this
algorithm, but we (regrettably) removed it to get under the page limit
and inadvertently removed it from the extended version at the same
time.  We will reinstate the full treatment in the extended report and
find a good way to summarize it in the final paper.

@@

The formal description is not very complex, but takes a good amount of
space to complete.  The formal description of RigidSynth first
describes an ordering on atoms, sequences, and DNF regular
expressions.  Next, this ordering is used to define a normalization
procedure for DNF regular expressions.  Finally, a procedure for
"zipping" two DNF regular expressions together is provided.  If these
two DNF regular expressions can be zipped, then there is a lens
between them.  A transformation from zipped DNF regular expressions
and the permutations that normalize them is then provided. These
formalizations are not very complex, but they take up a lot of space.

EVALUATION

A third area of concern was our evaluation, specifically the lack of
comparisons with other tools or a quantitative evaluation of the heuristics
we chose in searching for lenses.

Above, we discussed the fact that there are fundamental differences between
the kinds of transformations our tool can perform and the kinds of
transformations tools such as FlashFill and FlashExtract can perform.
We can quantify the set of benchmarks that each tool can process (at least
in one direction).

We agree that quantifying the impact of each heuristic we use would
strengthen our paper.  We will instrument our implementation to gather
data on the effectiveness of each heuristic individually.
Anecdotally, we invented each of these heuristics when more naive
approaches failed.  For example, reviewer D asks whether we "guess an
arbitrary regular expression to deal with composition."  Actually, our
first attempt at this problem tried to do that -- it did not use DNF
lenses and it could not synthesize anything beyond simple
microbenchmarks.  Moreover, this initial implementation did not even
try to search for programs that might use composition (a second issue
our revised system overcomes)!  Our priority queue and ranking system
were also motivated by failures endured by simpler system: those
simple systems were able to synthesize complicated lenses when the
source and target formats were similar, but not when the formats
differed significantly.

@@

BIJECTIVE LENS RESTRICTION

Bijective lenses are a subset of Boomerang's language of lenses, but
they are an important subset that both admit many useful
transformations and that also present some of the crucial technical
challenges.  In particular, overcoming the difficulties introduced by
regular expression typing is a fundamental challenge that is useful
for synthesizing larger subsets of Boomerang and any language related
to Boomerang.  (We also hope many of the principles introduced here
will apply to languages such as XDuce and other DTD or XML translation
languages, whose schema depend upon regular expressions.)

@@

===========================================================================
===========================================================================
The remainder of this document -- to be read only as needed -- responds in
detail to individual concerns not already addressed.

==========================
Review A:
==========================

- The search algorithm, particularly RigidSynth, is described almost
completely in prose, rather than with inductively defined rules.

As mentioned above, we have formalized this as a collection of
inference rules.  We regrettably omitted it for space reasons from
the conference submission.  We can include it in a future version.

- The synthesized lenses are not shown or discussed in much detail,
  nor are they compared to hand-written ones (when possible).  How do
  these compare to hand-written types/terms in Augeas? How do the
  inferred lenses compare to hand-written terms in Augeas (or Flash
  Fill, when appropriate)?

Our implementation puts some work into making the synthesized lenses
readable.  We perform the following transformations (among others):

@@

1. Trivial sublenses are removed
2. Sublenses expressing identity transformations are turned into identities
3. Identity transformations are combined when possible
4. Distributed transformations are factored when possible

The lens we presented above bibtex_to_endnote_au, is an autogenerated
lens.  Overall, we believe bibtex_to_endnote_au is quite similar to
what would be written in Augeas by hand.  One thing a human might
choose to do is to notice that:

lens_swap (NAME . del ",") (lens_swap WSP NAME)*

appears twice.  One could extract this common subexpression and bind it to a
name.  (We could implement a common sub-expression elimination pass but did 
not do so in the submission.)

@@

- How many of surveyed FlashFill examples were close to bijective? Were there
  only the 3 included in the evaluation?

Few of the surveyed FlashFill examples were "close to" bijective.  Most of
them are projections.  Bijections are typically used to synchronize two
data sources in different formats.  Hence the domains are different.  Our
purpose in considering the FlashFill examples was to illustrate that our
tool could handle the bijective data that FlashFill could handle.  And
as we mentioned earlier, there are many formats that we can handle that
FlashFill cannot.

- What do the type specifications look like?

Please scroll down to reviewer D (or grep for "cust/bib_prob" or
"aug/cron" below).  You will see that the formats we are describing
are really quite complicated.  Again, we would like to emphasize the
tradeoff we are studying vs a tool like FlashFill: we ask users to
supply information of kind they are familiar with (regular
expressions), which does clearly take work.  However, the payback is
that we now have a lot of leverage unavailable to other tools, and as
result, makes it possible for us to synthesize transformations over
very complex formats (with many nested iterations and unions that
other tools cannot handle).

- Is there reason to believe that the approach will or will not be
amenable for larger subsets of Boomerang?"

All of Boomerang is oriented around regular expression typing.
Showing how to manage the core problem of regular expression typing
and equivalence appears to be a crucial first step that any subsequent
tool will need to build on.  We are optimistic this foundation will
support larger subsets of Boomerang.

@@

==========================
Review B:
==========================

- I am not sure the remark about the implementation spanning '5515 lines' is as
positive as you make it sound at the beginning of Section 6. There is no
pseudocode for the main part of the algorithm, a function called RigidSynth.
This makes one wonder how complex it really is and, more worryingly, as a
result, there is also no correctness proof and no complexity analysis.

We regrettably omitted a formalization from the paper for space
reasons.  If authors are given additional pages in a final version of
the paper, we can include a version.  Much of the code is not for the
core synthesis algorithm, but for language transformations, pretty
printing, lens simplification, and DFA operations.  The portion of our
code corresponding to RigidSynth is just over 400 lines.

@@

- some proofs are missing

Please refer to our appendix, where we go into a detailed proof of these
theorems.  (The appendix was submitted with the paper as auxiliary materials ---
please ask other reviewers or the program chair if you have difficulty
locating the auxiliary files on hotcrp.)

@@

==========================
Review D:
==========================

- The algorithmic contributions (particularly, type-directedness) are often
  unclear.

The key algorithmic contribution is our technique for reducing the
language of general lenses to the language of DNF lenses.  We give
a formal completeness argument for this reduction.

After this reduction, our algorithm is has two phases, which alternate.
One phase rewrites regular expressions according to our rules.  It is
not type directed.  The second phase is type directed:  it is a
divide-and-conquer algorithm that makes a different decision based on
whether the top level constructor is concatenation, union or kleene star.

@@

      typedef PREFIX_DICT = "{\"prefix\"=" ("true" | "false") "}";;
      typedef TIME_DICT = "{\"minute\"=" NUMBER ",\"ws1\"=" INDENT_REQ ",\"hour\"=" NUMBER 
        ",\"ws2\"=" INDENT_REQ ",\"dayofmonth\"=" NUMBER ",\"ws3\"=" INDENT_REQ 
        ",\"month\"=" RANGE ",\"ws4\"=" INDENT_REQ ",\"dayofweek\"=" RANGE "}";;
      typedef SCHEDULE_DICT = "{\"schedule\"=\"" SCHEDULE_VALUE "\"}";;
      typedef ENTRY_DICT = "{\"indent\"=\"" INDENT "\"," PREFIX_DICT "," (TIME_DICT | SCHEDULE_DICT)
        ",\"indent2\"=\"" INDENT_REQ "\",\"user\"=\"" USER "\",\"indent3\"=\""
        INDENT_REQ "\",\"command\"=\"" SHELLCOMMAND "\"}";;
      typedef SHELL_DICT = "{\"varname\"=\"" SHELLVAR_NAME "\",\"value\"=\"" SHELLVALUE_NAME "\"}";;
      typedef CRON_DICT = ((EMPTYDICT | SHELL_DICT | COMMENT_DICT | ENTRY_DICT) "\n")*;;

      cron_lens = [CRON_DICT <=> CRON {}]

==========================
Review E:
==========================

- Why can’t the user provide regular expressions in first place that 
are aligned?

Describing a single format using a regular expression is a standard
task that programmers understand.  Aligning two regular expressions
is a level more difficult.  Because there are translations going on,
figuring out what needs to be aligned with what does not seem so easy.
Consider the two regular expressions (A|B)C* and
(A|C(C*)A)|(B|B(C*)C).  We believe it is quite 
nontrivial for a user to identify that
the equivalences that must be applied to reach two well aligned
lenses.  Furthermore, the aligned regular expressions could become
difficult to read and maintain: (A|B)(e+C+CC*C) is less understandable
than (A|B)C*.  Surely, avoiding this complexity is a win for the user.
Fortunately, we are successful in our efforts to do alignment
automatically so we can relieve programmers of this burden.

@@

- Is there a restriction that the user-provided regular expressions need to
  parse the source and target formats unambiguously? For example, if a user
  provides the regular expression for a format as Name* Name*, and the input is
  John Conway, would the system disallow such user inputs?

In general, the regular expressions need to be unambiguous.  If they
aren't then the kind of "round tripping" laws one would want simply
won't hold -- i.e. when you translate back-and-forth, you may wind up
corrupting the data.  (And making standard assumptions, like longest
prefix match does not work either).  Nate Foster covers this in depth
in his thesis.  Having said that, there are some places that ambiguity
is allowed.  In our system, we allow users to write ambiguous regular
expressions if they only want to perform the identity transformation
on that part of the string.  (If they want to carve up and rearrange
the data, the disambiguity constraints are important.)

@@

- The synthesis algorithm section presents several heuristics to assign scores
  and pseudometric scores to a pair of DNF regular expressions. How were these
  heuristics chosen? What is the effect of these heuristics on the benchmark
  problems?

Discussed above in the general section.

==========================
Review F:
==========================

- The examples from FlashFill and are those that can already be synthesized 
  using FlashFill by just providing input / output examples. Also, can the 
  authors describe how complex are the regular expressions that users need 
  to provide?

Please grep above for "aug/cron" and "cust/bib_prob" to see some example 
specs.

The purpose of the FlashFill examples was just to illustrate that our
system, with the extra information we supply it, certainly has no
problem with the FlashFill examples that are bijective.  As mentioned
above, most of the other benchmarks cannot be handled by FlashFill
because they contain nested iterations.  In total, 17 of the examples
contain nested iterations.  The most complex have 4 levels of nested
iterations.  These examples cannot be handled by FlashFill.

- Given that the same data format can be
  described using different semantically equivalent expressions, how do they
  affect the effectiveness of the synthesis process?

As we discuss in Speed of Synthesis, in the Evaluation section, the
difference in the presentation of the regular expressions can have an
impact on the performance of synthesis.  The "farther away" the two
pairs of expressions are, the more search is needed to find
sufficiently compatible pairs of regular expressions.  Much of the
focus of our heuristics is on identifying compatible regular
expression pairs during search.  Our experiments suggest they are
successful.

@@

- Finally, the real world contains many erroneous / dirty data. What happens if
  the user provide the incorrect regular expression or inconsistent input /
  output examples?

Briefly, one can test the regular expression against available data to
see if it parses the available data, but we offer no special support
here --- doing so is beyond the scope of this paper. Our experiments
suggest we need very few examples -- often zero or one.  If in the one
example the user puts the data in the wrong place, the user will
simply have specified the wrong transformation.  Developing some more
general machinery for analyzing many examples, some of which may be
noisy/buggy, is beyond the scope of this paper.

- Can the authors describe what kind of examples were needed to
  achieve the efficient synthesis results? In light of the "Importance of
  Examples" subsection on p.11, it seems that choosing "clever" examples is one
  of the key requirement for the tool to succeed in synthesizing a valid
  program.

"Clever" examples do not need to be provided to the system.  The "Avg # Exmpl"
column gets its value from randomly generated examples.

- What happens if multiple lens program are found for a given input?

The system returns the "simplest" program (we have a heuristic notion of
"simplest").

### 2017-01-29 paper/rebuttal.txt

TARGET AUDIENCE

Several reviewers asked who the intended users of this tool are and when the
use case is compelling.  The technology is aimed at programmers, and the
central idea is to exploit something that most programmers know well
(regular expressions) to help them write significant programs in a language
that they do not know well (lenses).

@@

EVALUATION

A third area of concern was our evaluation, specifically the lack of
comparisons with other tools or a quantitative evaluation of the heuristics
we chose in searching for lenses.

Above, we discussed the fact that there are fundamental differences between
the kinds of transformations our tool can perform and the kinds of
transformations tools such as FlashFill and FlashExtract can perform.
We can quantify the set of benchmarks that each tool can process (at least
in one direction).

We agree that quantifying the impact of each heuristic would strengthen the
paper.  We plan to instrument our implementation to gather data on the
effectiveness of each heuristic individually.

Anecdotally, we invented each of these heuristics when more naive approaches
failed.  For example, reviewer D asks whether we "guess an arbitrary regular
expression to deal with composition."  Actually, our first attempt at this
problem tried to do that -- it did not use DNF lenses and it could not
synthesize anything beyond simple microbenchmarks.  Moreover, this initial
implementation did not even try to search for programs that might use
composition (a second issue our revised system overcomes)!  The priority
queue and ranking system were also motivated by failures endured by earlier
implementations, which were able to synthesize complicated lenses when the
source and target formats were similar, but not when the formats differed
significantly.

@@

BIJECTIVE LENS RESTRICTION

Bijective lenses are a subset of Boomerang's language of lenses, but they
are an important subset that both admits many useful transformations and
also presents some of the crucial technical challenges.  In particular,
overcoming the difficulties introduced by regular expression typing is a
fundamental challenge that is useful for synthesizing larger subsets of
Boomerang and any language related to Boomerang.  We also hope many of the
principles introduced here will apply to languages such as XDuce and other
DTD or XML translation languages, whose schemas are based on upon regular
expressions.

@@

- Can the authors describe what kind of examples were needed to
  achieve the efficient synthesis results? In light of the "Importance of
  Examples" subsection on p.11, it seems that choosing "clever" examples is one
  of the key requirement for the tool to succeed in synthesizing a valid
  program.

"Clever" examples do not need to be provided to the system.  The "Avg # Exmpl"
column gets its value from randomly generated examples.

- What happens if multiple lens program are found for a given input?

The system returns the "simplest" program (using a heuristic notion of
"simplest").

### 2017-01-29 paper/rebuttal.txt

TARGET AUDIENCE

Several reviewers asked who the intended users of this tool are and when the
use case is compelling.  The technology is aimed at programmers, and the
central idea is to exploit something that most programmers know well
(regular expressions) to help them write significant programs in a language
that they do not know well (lenses).

Several reviewers wondered whether regular expressions were really necessary
at all and whether we could get away without requiring them.  One of the
reasons we feel our work is valuable is that it illustrates an important
practical tradeoff: by providing data formats, we can infer transformations
over significantly more sophisticated string data types than related systems
which do not use such information.  For instance, FlashFill cannot infer
nested iterations.  17 of our benchmarks have nested iterations and the
deepest nesting structure involved 4 levels of nesting.  As another example,
FlashExtract does not have unions, but it can represent them indirectly by
using options.  Still, in our tests (e.g. on the BibTeX data source),
FlashExtract could not infer transformations between string data types that
contain unions.  By the way, both FlashExtract and FlashFill are wonderful
systems, but they are optimized for other problems.  Our work demonstrates
the tradeoff between designs.  We will make this point more clearly in a
revised version of the paper, and in particular we will quantify the number
of benchmarks that we handle that cannot be handled by other string
synthesis systems.

### 2017-01-29 paper/rebuttal.txt

TARGET AUDIENCE

Several reviewers asked who the intended users of this tool are and
when the use case was compelling.  Briefly, the technology is aimed at
programmers.  If one were to ask random programmers whether they were
familiar with and knew how to use regular expressions, most would say
yes.  If one were to ask random programmers whether they were familiar
with lenses, 100% would say no.  The central idea is to exploit
something that most programmers know well (regular expressions) to
help them write significant programs in a language that they do not
know well (lenses).  Programmers are often reluctant to learn new
things.  Our system helps them avoid that.

Several reviewers wondered whether regular expressions were really
necessary at all and whether we could get away without requiring them.
One of the reasons we feel our work is valuable is that it illustrates
an important practical tradeoff: by providing data formats, we can
infer transformations over significantly more sophisticated string
data types than related systems which do not use such information.
For instance, FlashFill cannot infer nested iterations.  17 of our
benchmarks have nested iterations and the deepest nesting structure
involved 4 levels of nesting.  As another example, FlashExtract does
not have unions, but it can represent them indirectly by using
options.  Still, in our tests (e.g. on the BibTeX data source),
FlashExtract could not infer transformations between string data types
that contain unions. FlashExtract and FlashFill are wonderful systems,
but they are optimized for other problems.  Our work demonstrates the
tradeoff between designs.  We hope to make this point more clearly in
a revised version of the paper and to quantify the number of
benchmarks that we handle that cannot be handled by other string
synthesis systems.

@@

Several reviewers wondered whether lenses really were tricky to write
and suggested we show a concrete example (thank you -- we should
have).  The following lens translates between idealized BibTeX and
EndNote author formats (it was auto-generated by our tool):

  let bibtex_to_endnote_au
    : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
    del "author={"
      . ins "au - "
      . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
      . (del " and "
           . ins "\n au - "
           . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
      . del "}"

Writing this lens from scratch is quite tricky and fiddly.  The
trickiest part is the nested swaps, which are fiddly to think about
and get right.  We would also argue that the cognitive model overall
is tricky and quite different from what many programmers are used to
-- one has to think in two directions at once.  Importantly, this is
an incredibly simple example -- Augeas and Boomerang users regularly
work with lenses significantly larger than this.  For example,
Boomerang's full version of an idealized BibTeX converter is over 300
lines of code, and all but 4 of the Augeas programs we studied were
over 50 lines of code, with 4 over 100 lines of code.

@@

FORMALIZATION

A second major concern was the lack of a formal treatment of our
implementation, in particular the function RigidSynth.

In earlier drafts, we actually had a formal description of this
algorithm, but we (regrettably) removed it to get under the page limit
and inadvertently removed it from the extended version at the same
time.  We will reinstate the full treatment in the extended report and
find a good way to summarize it in the final paper.

The formal description is not very complex, but takes significant
space to complete.  The formal description of RigidSynth first
describes an ordering on atoms, sequences, and DNF regular
expressions.  Next, this ordering is used to define a normalization
procedure for DNF regular expressions.  Finally, a procedure for
"zipping" two DNF regular expressions together is provided.  If these
two DNF regular expressions can be zipped, then there is a lens
between them.  A transformation from zipped DNF regular expressions
and the permutations that normalize them is then provided. These
formalizations are not very complex, but they take up a lot of space.

@@

EVALUATION

A third area of concern was our evaluation, specifically the lack of
comparisons with other tools or a quantitative evaluation of the heuristics
we chose in searching for lenses.

Above, we discussed that there are fundamental differences between the
kinds of transformations our tool can perform and the kinds of
transformations tools such as FlashFill and FlashExtract can perform.
We can quantify the set of benchmarks that each tool can process (at
least in one direction).

@@

      cron_lens = [CRON_DICT <=> CRON {}]

==========================
Review E:
==========================

- Why can’t the user provide regular expressions in first place that 
are aligned?

Describing a single format using a regular expression is a standard
task that programmers understand.  Aligning two regular expressions is
considerably more difficult.  Due to the presence of translations,
figuring out what needs to be aligned with is no so easy.
Consider the two regular expressions (A|B)C* and
(A|C(C*)A)|(B|B(C*)C).  We believe it is quite nontrivial for a user
to identify which equivalences must be applied to reach two
well aligned lenses.  Furthermore, the aligned regular expressions
could become difficult to read and maintain: (A|B)(e+C+CC*C) is less
understandable than (A|B)C*.  Surely, avoiding this complexity is a
win for the user.  Fortunately, we are successful in our efforts to do
alignment automatically so we can relieve programmers of this burden.

### 2017-01-29 paper/rebuttal.txt

INTENDED USE CASE

Several reviewers asked who the intended users of this tool are and
when the use case is compelling.  The technology is aimed at
programmers, and the central idea is to exploit something that most
programmers know well (regular expressions) to help them write
significant programs in a language that they do not know well
(lenses).

Several reviewers wondered whether regular expressions were really
necessary at all and whether we could get away without requiring them.
One of the reasons we feel our work is valuable is that it illustrates
an important practical tradeoff: by providing data formats, we can
infer transformations over significantly more sophisticated string
data types than related systems which do not use such information.
For instance, FlashFill cannot infer nested iterations.  17 of our
benchmarks have nested iterations and the deepest nesting structure
involved 4 levels of nesting.  As another example, FlashExtract does
not have unions, but it can represent them indirectly by using
options.  Still, in our tests (e.g. on the BibTeX data source),
FlashExtract could not infer transformations between string data types
that contain unions. FlashExtract and FlashFill are wonderful systems,
but they are optimized for other problems.  Our work demonstrates the
tradeoff between designs.  We will make this point more clearly in a
revised version of the paper, and in particular we will quantify the
number of benchmarks that we handle that cannot be handled by other
string synthesis systems.

@@

Several reviewers wondered whether lenses really were tricky to write
and suggested we show a concrete example (thank you -- we should
have).  The following lens translates between idealized BibTeX and
EndNote author formats (it was auto-generated by our tool):

  let bibtex_to_endnote_au
    : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
    del "author={"
      . ins "au - "
      . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
      . (del " and "
           . ins "\n au - "
           . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
      . del "}"

Writing this lens from scratch is quite tricky and fiddly.  The
trickiest part is the nested swaps, which are fiddly to think about
and get right.  We would also argue that the cognitive model overall
is quite different from what many programmers are used to -- one has
to think in two directions at once.  Importantly, this is an
incredibly simple example -- Augeas and Boomerang users regularly work
with lenses significantly larger than this.  For example, Boomerang's
full version of an idealized BibTeX converter is over 300 lines of
code, and all but 4 of the Augeas programs we studied were over 50
lines of code, with 4 over 100 lines of code.

@@

FORMALIZATION

A second major concern was the lack of a formal treatment of our
implementation, in particular the function RigidSynth.

In earlier drafts, we actually had a formal description of this
algorithm, but we (regrettably) removed it to get under the page limit
and inadvertently removed it from the extended version at the same
time.  We will reinstate the full treatment in the extended report and
find a good way to summarize it in the final paper.

The formal description is not very complex, but takes significant
space to explain.  The formal description of RigidSynth first
describes an ordering on atoms, sequences, and DNF regular
expressions.  Next, this ordering is used to define a normalization
procedure for DNF regular expressions.  Finally, a procedure for
"zipping" two DNF regular expressions together is provided.  If 
two DNF regular expressions can be zipped, then there is a lens
between them.  A transformation from zipped DNF regular expressions
and the permutations that normalize them is then provided.

@@

A third area of concern was our evaluation, specifically the lack of
comparisons with other tools or a quantitative evaluation of the heuristics
we chose in searching for lenses.

Above, we discussed that there are fundamental differences between the
kinds of transformations our tool can perform and the kinds of
transformations tools such as FlashFill and FlashExtract can perform.
We can quantify the set of benchmarks that each tool can process (at
least in one direction).

We agree that quantifying the impact of each heuristic would
strengthen the paper.  We plan to instrument our implementation to
gather data on the effectiveness of each heuristic individually.

Anecdotally, we invented each of these heuristics when more naive
approaches failed.  For example, reviewer D asks whether we "guess an
arbitrary regular expression to deal with composition."  Actually, our
first attempt at this problem tried to do that -- it did not use DNF
lenses and it could not synthesize anything beyond simple
microbenchmarks.  Moreover, this initial implementation did not even
try to search for programs that might use composition (a second issue
our revised system overcomes)!  The priority queue and ranking system
were also motivated by failures encountered with earlier
implementations, which were able to synthesize complicated lenses when
the source and target formats were similar, but not when the formats
differed significantly.

### 2017-09-11 papers/BijectiveLensSynth/popl_18_rebuttal.txt

We thank the reviewers for their feedback.

...

Detailed Responses to individual questions by the reviewers follow below.

Review #18A
===========================================================================

Strengths/weaknesses

+ Interesting and potentially useful application of program synthesis to bidirectional programming
+ Nontrivial design/technical development
+ Approach implemented and evaluated on real(ish) examples
- Restriction of focus to bijective "lenses" not clearly justified
- Important parts of algorithm not explained (regarding how examples are used)
- Not clear how much work it is to adapt examples to be suitable for bijective lens synthesis, or how useful the results are w.r.t. the original task

But the main part of the paper is not self-contained - for example the
algorithm description relies on (example-dependent) orderings of sequences and atoms that
are not described; this is a key part of the algorithm since it is
where the provided examples are actually used.

I also have a concern: sicne the approach is
limited to bijective lenses, but many of the test examples were not
bijective, how much work was it to modify the examples
and how easy is it to manually hack the resulting bijective
lenses to make them work on the real data?  Is there any chance of
automating this or generalizing to handle non-bijective lenses
(perhaps with this algorithm as a subroutine)?

I could see future work building on this to handle the non-bijective case 
being a separate (and also interesting) paper.  But I would want the algorithm
to be described more clearly and more caveats about the examples.

p12 fig 6. I'm not sure how ule "DNF structural rewrite" could be
applied as the LHS uses the operators defined in fig. 5, rather than syntax

p17.  The desription of RigidSynth is incomplete.  It mentions orders
$\leq_{Seq}$ and $\leq_{Atom}$ (and variants that use the examples) that do
not appear to be defined.  What happens if there is ambiguity?  for
example what if we are trying to infer a lens of type (a | b | c) <->
(d|e|f) and we only have one example that says a <-> c?  If these
orderings are standard things then please give a reference.  If not,
they really need to be described in the paper.  The way examples are
used to align parts of regular expressions seems really critical to
this approach and is hardly explained.

p18: AllSome (alg. 3): I assume this is the obvious function of type
'a option list -> 'a list option (that is None if there is a None in
the list or Some [a1..an] if the list is [Some a1...Some an]) but
please clarify

p19: "easy to understand" - can you present any evidence for this?
Easy for whom?

p20: "We adapted these examples" please say mor about how many of the
examples needed to be adapted (all? a handful?) and how much work it
was, and how useful the resulting isomorphic lenses were w.r.t the
original data / how much additional work was needed to adapt them.

p20: "We developed a series of optimizations" - clarify that the
algorithm already described includes all of these, and what you are
going to do is disable them, rather than add more optimizations that
weren't discussed earlier.  Might be helpful to signpost what parts of
the algorithm are the "basic" algorithm and what parts are optional optimizations

p21: Fig 9 "Number of expansions" --> "Number of solutions"??

p21.  Also interesting to evaluate would be how time / number of
solutions varies (e.g. for the full algorithm) as the number of
examples provided varies.  Is the number of examples small? An active
approach where the algorithm asks the user for examples might also be
worth exploring (future work).

p24: "because our types do not have full canonical forms" - this is
the case for the equations presented in the paper, but what about
e.g. translating to automata and minimizing?

Review #18B
===========================================================================
Paper summary
-------------

However, I’m not sure what exact advantage the proposed approach brings to the
table. In particular, the user is still required to write exact input and output
types of the transformations via regular expressions. How hard is to write
lenses after one has such expressions? I’ll further detail my comparison now.
Also, the language has big limitations in terms of expressiveness.

### CONS:
- In the reviewer’s opinion, the specification mechanism is not very natural and
it’s unclear whether it actually simplifies writing bijective transformations.
This aspect is not evaluated and it would be nice to at least see the difference
in sizes between the domain/range regexes vs the Boomerang lenses. In general,
there is no report of the sizes of benchmarks.

- Very limited DSL for transformations. The proposed DSL can only describe
transformations that use identities and constant functions over input subparts.
For example, one cannot synthesize the function toUpperCase, which I imagine
could be useful for the kind of transformations appearing in the motivation.
Moreover, the functions can use each word in the input only once: one cannot
express the transformation abc->abcabc. Most bijective functions I can think of
(string encoders, compression algorithms, encryption algorithms) require complex
transformations over the input that go beyond identities. On this note, there
was a paper this year at PLDI on inverting string encoders (Automatic program
inversion using symbolic transducers, by Hu et al.), a topic which seems very
related, but for which I couldn’t find a citation.

- Somewhat limited evaluation. The evaluation is performed on two families of
benchmarks. The first family is an adaptation of format translation problems
from Augeas a tool for configuration editing and the second family is from
FlashFill a tool for synthesis of string transformations. All the benchmarks
presented are “adaptation” and it’s unclear whether the authors only picked
benchmarks for which they knew a priori a program in their DSL existed. Based
on the presented data, I cannot assess how often the presented technique could
be applied. Next, the comparison with the tool FlashFill/FlashExtract is somewhat
confusing. I suspect the authors had to provide some domain and range regular
expressions for the flash fill benchmarks, but this aspect, which is crucial to
help the search, is not mentioned. What happens if someone provides Sigma* and
Sigma* as domains and ranges to Optician. Also, a better explanation of the
difference between the DSLs of Lenses and Flash fill is required. For example,
FlashFill can reverse a string, but the proposed language can’t.

### QUESTIONS FOR REBUTTAL
- What is the difference in size between the regexes for domain/range spec and
the full lenses?
- What is the difference in size between manually written lenses and synthesized
ones?
- Clarify how the benchmarks were selected and whether any were not selected a
priori. Especially the flash* ones. Did you omit those that used character
transformations such as upper-case?
- Can the proposed DSL express any transformation that FlashFill can’t and
vice-versa?
- P19: maximally factors the concats and ors. What does this mean? Regex
minimization is a hard problem, what is that one factors exactly? I think it
would help to show one example of fully synthesized lens in the paper.

Minor:
- Eval the text says Full synthesized 49 benchmarks, but the plot in fig 8 only
shows 38/39?

Review #18C
===========================================================================

Overall merit
-------------
3. Weak reject - will not argue against

Reviewer expertise
------------------
Z. Outsider

Paper summary
-------------
Overall, this paper looks like a solid and useful contribution to the lens
literature. However, this reviewer has never used, let alone designed, a lens
library. As such, I found the paper very hard to read and assess. I would have
liked to see a clear problem statement (such as an example where it's annoying
to write the Boomerang code but easy to synthesize the lens using your approach
- why is there no Boomerang code for your example in Sec. 3?). It was also not
very clear to me what subset of Boomerang your approach supports. The
introduction says it is a "useful subset of Boomerang" - but what does that
mean?

I was also a bit confused about the preliminaries in Sec. 2. In particular, it
is not clear whether all or most of those preliminaries are from related work
(but there are no references) or whether they are contributions of this paper.

Minor comments:

- p.4, "naively using it in the context of lens synthesis presents several
problems"? Which problems? At this point I have no clue what "using it" may
even mean. The "big picture" is not clear at all.

- p. 3, "We have ommitted these theorems for space..." - maybe you mean
"proofs"?

-p. 7, "instead of maintaining server code for each endpoint, we envision..."
- but surely the changes to an API involve more than just data format changes.
For instance, single calls may have to be split into multiple calls. Hence it
does not seem to be realistic to maintain the old API using _only_ synthesized
lenses.

### 2017-09-11 papers/BijectiveLensSynth/popl_rebuttal_strategy_notes.txt

Summarize points they all raise, and questions for rebuttal:

Quotient lenses allow for removing whitespace

Reviewer 1:
Specify and talk about QREs, transformations can be factored into 3 parts,
and we have a way of specifying 2 of those parts.  1 part that transforms
data into a canonical form, transformations into canonical forms and then
same transforms as usual in the middle.
We've experimented with this and our current inference algorithm works
off the shelf with it.

Rebuttal to reviewer B issues is that:
Are there any xforms that we can synth and flashfill can't
Start with that..., talk about nested iterations etc.
One of largest strengths is that it can synthesize really complex ones that
can't be synthesized by FF

Point to thing thing in Excel, as stated it can't synthesize nested stars

The DSL allows for nested stars, but it doesn't seem to work

Copy & paste text of their questions,
Differences between FF and Optician
How do we transform the examples, describe changes
Any chance of automating this or handling, cut & paste question

	2 major classes of changes we had to make: 1 is whitespace, other is
 	  projection of data
	Bijections have to preserve the amount of whitespace, this can be
	  automated by putting a wrapper around the xform that normalizes
	  the whitespace
	Our current system does that (this is called a quotient lens)
	Second transformation is forgetting information, and then restoring
	  after a round-trip.  We again believe the core of our algorithm
	  can be reused for this, and we are working on this still, but it's
	  future work.  Say how we modified the examples - added in the
	  extraneous data.  Again, a wrapper that deleted that and restored
	  it on the reverse transformation could handle it and is not hard
	  to write - believe can be automated, but need to do the research.

Reviewers are worried we're cheating about experiments

	Try and avoid subjective wording
	Concrete facts about what's going on
	In whitespace, we're making sure that each side has same amount of
	  whitespace
